Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-02-05 causal-bandit/icml-2015/tech.tex
Tweak wording
many minor improvements and checking Mark's comments

\begin{proof}
Let $F$ be the event that there exists and $1 \leq i \leq N$ for which
\eq{
\left|\hat q_i - q_i\right| \geq \sqrt{\frac{6q_i}{T} \log \frac{2N}{\delta}}\,.
}
Then by the union bound and Lemma \ref{lem:conc1} we have $\P{F} \leq \delta$. The result will be completed by showing that
when $F$ does not hold we have $2m(\vec{q})/3 \leq m(\vec{\hat q}) \leq 2m(\vec{q})$.
Towards this end recall that $m(\vec{q}) = \min\set{i : q_i \geq 1 / i}$.
Therefore for $i > m$ we have $q_i \geq q_m \geq 1/m$ and so by Lemma \ref{lem:conc1} we have
\eq{
\frac{3}{4} 
&\geq \frac{1}{2} + \sqrt{\frac{3}{T} \log \frac{2N}{\delta}} 
\geq q_i + \sqrt{\frac{6q_i}{T} \log \frac{2N}{\delta}} 
\geq \hat q_i \\
&\geq q_i - \sqrt{\frac{6q_i}{T} \log \frac{2N}{\delta}}
\geq q_i - \sqrt{\frac{q_i}{8m}}
\geq \frac{1}{2m}\,.
}
Therefore by the pigeonhole principle we have $m(\vec{\hat q}) \leq 2m$.
For the other direction we proceed in a similar fashion. Since the failure event $F$ does not hold we have for $i \leq m$ that
\eq{
\hat q_i 
\leq q_i + \sqrt{\frac{6q_i}{T} \log\frac{2N}{\delta}} 
\leq \frac{1}{m} \left(1 + \sqrt{\frac{1}{8}}\right)
\leq \frac{3}{2m}\,.
}
Therefore $m(\vec{\hat q}) \geq 2m(\vec{q}) / 3$ as required. 
\end{proof}

@@

Assume without loss of generality that $q_1 \leq q_2 \leq \ldots \leq q_N \leq 1/2$.
For each $i$ define reward function $r_i$ by
\eq{
r_0(\boldsymbol{X}) &= \frac{1}{2} &
r_i(\boldsymbol{X}) &= \begin{cases}
\frac{1}{2} + \epsilon & \text{if } X_i = 1 \\
\frac{1}{2} & \text{otherwise}\,,
\end{cases}
}
where $1/4 \geq \epsilon > 0$ is some constant to be chosen later.
We abbreviate $R_{T,i}$ to be the expected simple regret incurred when interacting with the
environment determined by $\boldsymbol{q}$ and $r_i$. Let $\operatorname{P}_i$ be the corresponding measure
on all observations over all $T$ rounds and $\EE_i$ the expectation with respect to $\operatorname{P}_i$. By Lemma 2.6 by \cite{Tsy08} we have
\eq{
\Prz{\hat a^*_T = a^*} + \Pri{\hat a^*_T \neq a^*} \geq \exp\left(-\KL(\operatorname{P}_0, \operatorname{P}_i)\right)\,,
}
where $\KL(\Ps_0, \Ps_i)$ is the KL divergence between measures $\operatorname{P}_0$ and $\operatorname{P}_i$.
Let $T_i(T) = \sum_{t=1}^T \ind{a_t = do(X_i = 1)}$ be the total number of times the learner intervenes on variable $i$ by setting it to $1$.
Then for $i \leq m$ we have $q_i \leq 1/m$ and the KL divergence between $\Ps_0$ and $\Ps_i$ may be bounded using the telescoping property (chain rule) and
by bounding the local KL divergence by the $\chi$-squared distance as by \cite{Auer1995}. This leads to 
\eq{
\KL(\Ps_0, \Ps_i) 
&\leq 6\epsilon^2 \EE_0\left[\sum_{t=1}^T \ind{X_{t,i} = 1}\right] 
\leq 6\epsilon^2 \left(\EE_0 T_i(T) + q_i T\right) 
\leq 6\epsilon^2 \left(\EE_0 T_i(T) + \frac{T}{m}\right)\,.
}
Define set $A = \set{i \leq m : \EE_0 T_i(T) \leq 2T / m}$.
Then for $i \in A$ and choosing $\epsilon = \min\set{1/4, \sqrt{m/(18T)}}$ we have
\eq{
\KL(\Ps_0, \Ps_i) \leq \frac{18T\epsilon^2}{m} = 1\,. 
}
Now $\sum_{i=1}^m \EE_0 T_i(T) \leq T$, which implies that $|A| \geq m/2$.
Therefore
\eq{
\sum_{i \in A} \Pri{I_T \neq i} 
\geq \sum_{i \in A} \exp\left(-\KL(\Ps_0, \Ps_i)\right) - 1
\geq \frac{|A|}{e} - 1 
\geq \frac{m}{2e} - 1\,.
}
Therefore there exists an $i \in A$ such that
$\Pri{I_T \neq i} \geq \frac{\frac{m}{2e} - 1}{m}$. 
Therefore if $\epsilon < 1/4$ we have
\eq{
R_{T,i} \geq \frac{1}{2} \Pn{i}{\hat a^*_T \neq a^*} \epsilon \geq \frac{\frac{m}{2e} - 1}{2m} \sqrt{\frac{m}{18T}}\,.
}
Otherwise $m \geq 18T$ so $\sqrt{m/T} = \Omega(1)$ and
\eq{
R_{T,i} \geq \frac{1}{2} \Pn{i}{\hat a^*_T \neq a^*} \epsilon \geq \frac{1}{4} \frac{\frac{m}{2e} - 1}{2m} \in \Omega(1) 
}
as required.

>>>>

\begin{proof}
Let $F$ be the event that there exists and $1 \leq i \leq N$ for which
\eq{
\left|\hat q_i - q_i\right| \geq \sqrt{\frac{6q_i}{T} \log \frac{2N}{\delta}}\,.
}
Then by the union bound and Lemma \ref{lem:conc1} we have $\P{F} \leq \delta$. The result will be completed by showing that
when $F$ does not hold we have $2m(\vec{q})/3 \leq m(\vec{\hat q}) \leq 2m(\vec{q})$.
From the definition of $m(\vec{q})$ and our assumption on $\vec{q}$ we have for $i > m$ that $q_i \geq q_m \geq 1/m$ and so by Lemma \ref{lem:conc1} we have
\eq{
\frac{3}{4} 
&\geq \frac{1}{2} + \sqrt{\frac{3}{T} \log \frac{2N}{\delta}} 
\geq q_i + \sqrt{\frac{6q_i}{T} \log \frac{2N}{\delta}} 
\geq \hat q_i \\
&\geq q_i - \sqrt{\frac{6q_i}{T} \log \frac{2N}{\delta}}
\geq q_i - \sqrt{\frac{q_i}{8m}}
\geq \frac{1}{2m}\,.
}
Therefore by the pigeonhole principle we have $m(\vec{\hat q}) \leq 2m$.
For the other direction we proceed in a similar fashion. Since the failure event $F$ does not hold we have for $i \leq m$ that
\eq{
\hat q_i 
\leq q_i + \sqrt{\frac{6q_i}{T} \log\frac{2N}{\delta}} 
\leq \frac{1}{m} \left(1 + \sqrt{\frac{1}{8}}\right)
\leq \frac{3}{2m}\,.
}
Therefore $m(\vec{\hat q}) \geq 2m(\vec{q}) / 3$ as required. 
\end{proof}

@@

We follow a relatively standard path by choosing multiple environments that have different optimal arms, but which cannot all be statistically
separated in $T$ rounds.
Assume without loss of generality that $q_1 \leq q_2 \leq \ldots \leq q_N \leq 1/2$.
For each $i$ define reward function $r_i$ by
\eq{
r_0(\boldsymbol{X}) &= \frac{1}{2} &
r_i(\boldsymbol{X}) &= \begin{cases}
\frac{1}{2} + \epsilon & \text{if } X_i = 1 \\
\frac{1}{2} & \text{otherwise}\,,
\end{cases}
}
where $1/4 \geq \epsilon > 0$ is some constant to be chosen later.
We abbreviate $R_{T,i}$ to be the expected simple regret incurred when interacting with the
environment determined by $\boldsymbol{q}$ and $r_i$. Let $\operatorname{P}_i$ be the corresponding measure
on all observations over all $T$ rounds and $\EE_i$ the expectation with respect to $\operatorname{P}_i$. By Lemma 2.6 by \citet{Tsy08} we have
\eq{
\Prz{\hat a^*_T = a^*} + \Pri{\hat a^*_T \neq a^*} \geq \exp\left(-\KL(\operatorname{P}_0, \operatorname{P}_i)\right)\,,
}
where $\KL(\Ps_0, \Ps_i)$ is the KL divergence between measures $\operatorname{P}_0$ and $\operatorname{P}_i$.
Let $T_i(T) = \sum_{t=1}^T \ind{a_t = do(X_i = 1)}$ be the total number of times the learner intervenes on variable $i$ by setting it to $1$.
Then for $i \leq m$ we have $q_i \leq 1/m$ and the KL divergence between $\Ps_0$ and $\Ps_i$ may be bounded using the telescoping property (chain rule) and
by bounding the local KL divergence by the $\chi$-squared distance as by \citet{Auer1995}. This leads to 
\eq{
\KL(\Ps_0, \Ps_i) 
&\leq 6\epsilon^2 \EE_0\left[\sum_{t=1}^T \ind{X_{t,i} = 1}\right] 
\leq 6\epsilon^2 \left(\EE_0 T_i(T) + q_i T\right) 
\leq 6\epsilon^2 \left(\EE_0 T_i(T) + \frac{T}{m}\right)\,.
}
Define set $A = \set{i \leq m : \EE_0 T_i(T) \leq 2T / m}$.
Then for $i \in A$ and choosing $\epsilon = \min\set{1/4, \sqrt{m/(18T)}}$ we have
\eq{
\KL(\Ps_0, \Ps_i) \leq \frac{18T\epsilon^2}{m} = 1\,. 
}
Now $\sum_{i=1}^m \EE_0 T_i(T) \leq T$, which implies that $|A| \geq m/2$.
Therefore
\eq{
\sum_{i \in A} \Pri{\hat a^*_T \neq a} 
\geq \sum_{i \in A} \exp\left(-\KL(\Ps_0, \Ps_i)\right) - 1
\geq \frac{|A|}{e} - 1 
\geq \frac{m}{2e} - 1\,.
}
Therefore there exists an $i \in A$ such that
$\Pri{\hat a^*_T \neq a^*} \geq \frac{\frac{m}{2e} - 1}{m}$. 
Therefore if $\epsilon < 1/4$ we have
\eq{
R_{T,i} \geq \frac{1}{2} \Pn{i}{\hat a^*_T \neq a^*} \epsilon \geq \frac{\frac{m}{2e} - 1}{2m} \sqrt{\frac{m}{18T}}\,.
}
Otherwise $m \geq 18T$ so $\sqrt{m/T} = \Omega(1)$ and
\eq{
R_{T,i} \geq \frac{1}{2} \Pn{i}{\hat a^*_T \neq a^*} \epsilon \geq \frac{1}{4} \frac{\frac{m}{2e} - 1}{2m} \in \Omega(1) 
}
as required.
